---
title: "Predict food trend using social media data"
output: html_document
---

Objective is to predict total sales volume using social media data. 


```{r}
require(scales)
```


Read datasets for clementine sales and social media.  

```{r}
social <- read.csv("/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clementine_social.csv")

# Check summary statistics
#summary(social)
head(social[1:9])

sales <- read.csv("/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clementine_sales.csv")

# Check summary statistics
#summary(sales)
head(sales)
```

Let's make some time series plots for social media mentions and sales data for Clementines. 

```{r}
# Social media mentions plot

# Date formatting 
analysis_date = as.Date(social[,1], "%m/%d/%y")
str(analysis_date)

# Create a variable for month to aggregate / sum all observations for a particular month 
social$month = as.Date(cut(analysis_date, breaks = "month"))
#social

total_mentions = social[,9]

# Create a dataset containing analysis date, month of analysis date and total social media mentions for plotting a observations aggregated at a monthly level. 
social1 <- data.frame(analysis_date, social$month, total_mentions)
#social1

require(ggplot2)
# Time series plot by monthly total observations
p <- ggplot(social1, aes(x=social.month, y=total_mentions)) 
p + geom_line() 
# add up all observations(total_mentions) for the month
p = p + stat_summary(fun.y = sum, geom = "line")
p = p + ggtitle ("Total clementine mentions across different social media outlets") 
p + xlab("Date") +  ylab ("Total mentions for Clementine")
#p + scale_x_date(labels = date_format("%y/%m"), breaks = "1 month")

# Plotting time series with curve / trendline 
#ggplot(socialts,aes(x=social.month, y=total_mentions)) + 
#  geom_point(aes(colour = total_mentions)) +
#  scale_colour_gradient2(low = "blue", mid = "green" , high = "red", midpoint = 4738 ) + 
#  geom_smooth(color = "red",size = 1) +
#  scale_y_continuous(limits = c(899,23695)) +
#  ggtitle ("Total clementine mentions across different social media outlets") +
#  xlab("Date") +  ylab ("Total mentions for Clementine")
   

```

```{r}
# Sales data plot 

# Date formatting 
period = as.Date(sales[,1], "%m/%d/%y")
str(period)

Volume = sales[,6]

# Create a variable for month to aggregate / sum all observations for a particular month 
sales$month = as.Date(cut(period, breaks = "month"))
#sales

# Bind period, sales month and volume into a data frame 
sales1 <- data.frame(period, sales$month, Volume)

# Now, let's plot time series of montly aggregated total Clementine sales volume
p <- ggplot(sales1, aes(x=sales.month, y=Volume))
p + geom_line() 
# add up all observations(total_mentions) for the month
p = p + stat_summary(fun.y = sum, geom = "line")
#p = p + scale_y_continuous(limits = c(211151,10017414))
p = p + ggtitle ("Total volume of clementine sales (in U.S)") 
p + xlab("Date") +  ylab ("Total Volume of Clementine sales")
```





Remove seasonal component. The remaining time series should be trend and random noise. 

Why do we seasonally adjust data? 

http://www.bls.gov/cps/seasfaq.htm 

In order to obtain / understand the true underlying trend, we need to seasonally adjust the data. Upon, seasonally adjusting the data we get the trend component, seasonal component and irregular time series component. 

Decomposing the time series into trend, seasonal and irregular components. 

```{r}
# Columns required time and volume

#Using zoo object
require(zoo)
salests = sales[, c(1,6)]
salestsz = read.zoo(salests, format = "%m/%d/%y", index.column = 1)
View(salestsz)


#Decompose zoo object
salestszDec <- ts(salestsz, frequency=52, start=c(2010, 1))
View(salestszDec)
salestsDec1 <- decompose(salestszDec, type=c("additive"))
View(salestszDec)
plot(salestsDec1)


# Frequency = 52 weeks in year. i.e 52 # of observations in one year. 
#salestsDec <- ts(salests[,2], frequency=52, start=c(2010, 1), end=c(2014,12))
#salestsDec <- decompose(salestsDec, type=c("additive"))
#plot(salestsDec)

```

Plot seasonally adjusted time series 

```{r}
# Check object type
class(salestsDec1$seasonal)
class(salestszDec)
# Both are 'ts' object, so we need to convert to numeric to be able to perform math operation of subtracting seasonal component from the original series. 


salestszDec <- as.numeric(unlist(salestszDec))
class(salestszDec)
salestszSeasonallyAdjusted <- salestszDec - salestsDec1$seasonal

#Plot seasonally adjusted time series
plot(salestszSeasonallyAdjusted, xlab="Time", ylab="Volume of sales", main="Seasonally adjusted clementine sales volume")

clemSalesFinal <- cbind(index(salestsz),salestsDec1$x,salestsDec1$seasonal,salestszSeasonallyAdjusted)

require(xlsx)
# Don't re-run, the file will append the current combined file. 
write.xlsx(clemSalesFinal, file="/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clem_sales_seasonally_adjusted_1.xlsx", sheetName="sheet1", col.names=TRUE, row.names=FALSE, append=FALSE, showNA=TRUE)

index(salestsz)
```


Decompose the time series and adjust seasonality for $$ amounts from clementine sales. 

```{r}
# Columns required time and Price

#Using zoo object for time indexing
require(zoo)
clemSalesPrice = sales[, c(1,5)]
clemSalesPricez = read.zoo(clemSalesPrice, format = "%m/%d/%y", index.column = 1)


#Decompose zoo object
clemSalesPricezDec <- ts(clemSalesPricez, frequency=52, start=c(2010, 1))
clemSalesPriceDec1 <- decompose(clemSalesPricezDec, type=c("additive"))
plot(clemSalesPriceDec1)


# Frequency = 52 weeks in year. i.e 52 # of observations in one year. 
#salestsDec <- ts(salests[,2], frequency=52, start=c(2010, 1), end=c(2014,12))
#salestsDec <- decompose(salestsDec, type=c("additive"))
#plot(salestsDec)

```

Plot seasonally adjusted time series (Price)

```{r}
# Check object type
class(clemSalesPriceDec1$seasonal)
class(clemSalesPricezDec)
# Both are 'ts' object, so we need to convert to numeric to be able to perform math operation of subtracting seasonal component from the original series. 


clemSalesPricezDec <- as.numeric(unlist(clemSalesPricezDec))
class(clemSalesPricezDec)
clemSalesPriceSeasonallyAdjusted <- clemSalesPricezDec - clemSalesPriceDec1$seasonal

#Plot seasonally adjusted time series
plot(clemSalesPriceSeasonallyAdjusted, xlab="Time", ylab="Dollar amount", main="Seasonally adjusted clementine sales (Dollar amount)")

# Combine columns 
clemsalesPriceFinal <- cbind(clemSalesPriceDec1$x,clemSalesPriceDec1$seasonal,clemSalesPriceSeasonallyAdjusted)


require(xlsx)
# Don't re-run, the file will append the current combined file. Write original price, seasonal component and seasonally adjusted values to en excel file

write.xlsx(clemsalesPriceFinal, file="/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clem_sales_seasonally_adjusted_price.xlsx", sheetName="sheet1", col.names=TRUE, row.names=FALSE, append=FALSE, showNA=TRUE)
```


Decompose the time series and adjust seasonality for social media mentions.

Twitter mentions

```{r}
# Columns required time and twitter mentions

#Using zoo object for time indexing
require(zoo)
clemTwitter = social[, c(1,6)]
clemTwitterz = read.zoo(clemTwitter, format = "%m/%d/%y", index.column = 1)
View(clemSalesPricez)


#Decompose zoo object
clemTwitterzDec <- ts(clemTwitterz, frequency=52, start=c(2010, 1))
clemTwitterDec1 <- decompose(clemTwitterzDec, type=c("additive"))
plot(clemTwitterDec1)


# Frequency = 52 weeks in year. i.e 52 # of observations in one year. 
#salestsDec <- ts(salests[,2], frequency=52, start=c(2010, 1), end=c(2014,12))
#salestsDec <- decompose(salestsDec, type=c("additive"))
#plot(salestsDec)

```

Plot seasonally adjusted time series (Twitter)

```{r}
# Check object type
class(clemTwitterDec1$seasonal)
class(clemTwitterzDec)
# Both are 'ts' object, so we need to convert to numeric to be able to perform math operation of subtracting seasonal component from the original series. 


clemTwitterzDec <- as.numeric(unlist(clemTwitterzDec))
class(clemTwitterzDec)
clemTwitterSeasonallyAdjusted <- clemTwitterzDec - clemTwitterDec1$seasonal

# Plot seasonally adjusted time series
plot(clemTwitterSeasonallyAdjusted, xlab="Time", ylab="Twitter mentions", main="Seasonally adjusted clementine twitter mentions")

# Combine columns 
clemTwitterFinal <- cbind(clemTwitterDec1$x,clemTwitterDec1$seasonal,clemTwitterSeasonallyAdjusted)

require(xlsx)
# Don't re-run, the file will append the current combined file. Write original price, seasonal component and seasonally adjusted values to en excel file

write.xlsx(clemTwitterFinal, file="/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clem_social_seasonally_adjusted_twitter.xlsx", sheetName="sheet1", col.names=TRUE, row.names=FALSE, append=FALSE, showNA=TRUE)

```


Facebook mentions

```{r}
# Columns required time and facebook mentions

#Using zoo object for time indexing
require(zoo)
clemFacebook = social[, c(1,5)]
clemFacebookz = read.zoo(clemFacebook, format = "%m/%d/%y", index.column = 1)
View(clemFacebookz)


#Decompose zoo object
clemFacebookzDec <- ts(clemFacebookz, frequency=52, start=c(2010, 1))
clemFacebookDec1 <- decompose(clemFacebookzDec, type=c("additive"))
plot(clemFacebookDec1)


# Frequency = 52 weeks in year. i.e 52 # of observations in one year. 
#salestsDec <- ts(salests[,2], frequency=52, start=c(2010, 1), end=c(2014,12))
#salestsDec <- decompose(salestsDec, type=c("additive"))
#plot(salestsDec)

```

Plot seasonally adjusted time series (Facebook)

```{r}
# Check object type
class(clemFacebookDec1$seasonal)
class(clemFacebookzDec)
# Both are 'ts' object, so we need to convert to numeric to be able to perform math operation of subtracting seasonal component from the original series. 


clemFacebookzDec <- as.numeric(unlist(clemFacebookzDec))
class(clemFacebookzDec)
clemFacebookSeasonallyAdjusted <- clemFacebookzDec - clemFacebookDec1$seasonal

#Plot seasonally adjusted time series
plot(clemFacebookSeasonallyAdjusted, xlab="Time", ylab="Facebook mentions", main="Seasonally adjusted clementine Facebook mentions")

#Combine data
clemFacebookFinal <- cbind(clemFacebookDec1$x,clemFacebookDec1$seasonal,clemFacebookSeasonallyAdjusted)

require(xlsx)
# Don't re-run, the file will append the current combined file. Write original price, seasonal component and seasonally adjusted values to en excel file

write.xlsx(clemFacebookFinal, file="/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clem_social_seasonally_adjusted_fb.xlsx", sheetName="sheet1", col.names=TRUE, row.names=FALSE, append=FALSE, showNA=TRUE)

```

Blogs mentions

```{r}
# Columns required time and blog mentions

#Using zoo object for time indexing
require(zoo)
clemBlog = social[, c(1,2)]
clemBlogz = read.zoo(clemBlog, format = "%m/%d/%y", index.column = 1)
View(clemBlogz)


#Decompose zoo object
clemBlogzDec <- ts(clemBlogz, frequency=52, start=c(2010, 1))
clemBlogDec1 <- decompose(clemBlogzDec, type=c("additive"))
plot(clemBlogDec1)


# Frequency = 52 weeks in year. i.e 52 # of observations in one year. 
#salestsDec <- ts(salests[,2], frequency=52, start=c(2010, 1), end=c(2014,12))
#salestsDec <- decompose(salestsDec, type=c("additive"))
#plot(salestsDec)

```

Plot seasonally adjusted time series (Blog)

```{r}
# Check object type
class(clemBlogDec1$seasonal)
class(clemBlogzDec)
# Both are 'ts' object, so we need to convert to numeric to be able to perform math operation of subtracting seasonal component from the original series. 


clemBlogzDec <- as.numeric(unlist(clemBlogzDec))
class(clemBlogzDec)
clemBlogSeasonallyAdjusted <- clemBlogzDec - clemBlogDec1$seasonal

#Plot seasonally adjusted time series
plot(clemBlogSeasonallyAdjusted, xlab="Time", ylab="Blog mentions", main="Seasonally adjusted clementine Blog mentions")

# Combine data 
clemBlogFinal <- cbind(clemBlogDec1$x,clemBlogDec1$seasonal,clemBlogSeasonallyAdjusted)


require(xlsx)
# Don't re-run, the file will append the current combined file. Write original price, seasonal component and seasonally adjusted values to en excel file

write.xlsx(clemBlogFinal, file="/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clem_social_seasonally_adjusted_blog.xlsx", sheetName="sheet1", col.names=TRUE, row.names=FALSE, append=FALSE, showNA=TRUE)

```

Forums mentions

```{r}
# Columns required time and forums mentions

#Using zoo object for time indexing
require(zoo)
clemForums = social[, c(1,3)]
clemForumsz = read.zoo(clemForums, format = "%m/%d/%y", index.column = 1)
View(clemForumsz)


#Decompose zoo object
clemForumszDec <- ts(clemForumsz, frequency=52, start=c(2010, 1))
clemForumsDec1 <- decompose(clemForumszDec, type=c("additive"))
plot(clemForumsDec1)


# Frequency = 52 weeks in year. i.e 52 # of observations in one year. 
#salestsDec <- ts(salests[,2], frequency=52, start=c(2010, 1), end=c(2014,12))
#salestsDec <- decompose(salestsDec, type=c("additive"))
#plot(salestsDec)

```

Plot seasonally adjusted time series (Forums)

```{r}
# Check object type
class(clemForumsDec1$seasonal)
class(clemForumszDec)
# Both are 'ts' object, so we need to convert to numeric to be able to perform math operation of subtracting seasonal component from the original series. 


clemForumszDec <- as.numeric(unlist(clemForumszDec))
class(clemForumszDec)
clemForumsSeasonallyAdjusted <- clemForumszDec - clemForumsDec1$seasonal

#Plot seasonally adjusted time series
plot(clemForumsSeasonallyAdjusted, xlab="Time", ylab="Forum mentions", main="Seasonally adjusted clementine Forum mentions")

#Combine data
clemForumsFinal <- cbind(clemForumsDec1$x,clemForumsDec1$seasonal,clemForumsSeasonallyAdjusted)
require(xlsx)
# Don't re-run, the file will append the current combined file. Write original price, seasonal component and seasonally adjusted values to en excel file

write.xlsx(clemForumsFinal, file="/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clem_social_seasonally_adjusted_forums.xlsx", sheetName="sheet1", col.names=TRUE, row.names=FALSE, append=FALSE, showNA=TRUE)

```

Reviews mention

```{r}
# Columns required time and blog mentions

#Using zoo object for time indexing
require(zoo)
clemReviews = social[, c(1,4)]
clemReviewsz = read.zoo(clemReviews, format = "%m/%d/%y", index.column = 1)
View(clemReviewsz)


#Decompose zoo object
clemReviewszDec <- ts(clemReviewsz, frequency=52, start=c(2010, 1))
clemReviewsDec1 <- decompose(clemReviewszDec, type=c("additive"))
plot(clemReviewsDec1)


# Frequency = 52 weeks in year. i.e 52 # of observations in one year. 
#salestsDec <- ts(salests[,2], frequency=52, start=c(2010, 1), end=c(2014,12))
#salestsDec <- decompose(salestsDec, type=c("additive"))
#plot(salestsDec)

```

Plot seasonally adjusted time series (Reviews)

```{r}
# Check object type
class(clemReviewsDec1$seasonal)
class(clemReviewszDec)
# Both are 'ts' object, so we need to convert to numeric to be able to perform math operation of subtracting seasonal component from the original series. 


clemReviewszDec <- as.numeric(unlist(clemReviewszDec))
class(clemReviewszDec)
clemReviewsSeasonallyAdjusted <- clemReviewszDec - clemReviewsDec1$seasonal

#Plot seasonally adjusted time series
plot(clemReviewsSeasonallyAdjusted, xlab="Time", ylab="Reviews mentions", main="Seasonally adjusted clementine Reviews mentions")

#Combine data
clemReviewsFinal <- cbind(clemReviewsDec1$x,clemReviewsDec1$seasonal,clemReviewsSeasonallyAdjusted)

require(xlsx)
# Don't re-run, the file will append the current combined file. Write original price, seasonal component and seasonally adjusted values to en excel file

write.xlsx(clemReviewsFinal, file="/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clem_social_seasonally_adjusted_reviews.xlsx", sheetName="sheet1", col.names=TRUE, row.names=FALSE, append=FALSE, showNA=TRUE)

```

Comments mention 

```{r}
# Columns required time and comments mentions

#Using zoo object for time indexing
require(zoo)
clemComments = social[, c(1,7)]
clemCommentsz = read.zoo(clemComments, format = "%m/%d/%y", index.column = 1)
View(clemCommentsz)


#Decompose zoo object
clemCommentszDec <- ts(clemCommentsz, frequency=52, start=c(2010, 1))
clemCommentsDec1 <- decompose(clemCommentszDec, type=c("additive"))
plot(clemCommentsDec1)


# Frequency = 52 weeks in year. i.e 52 # of observations in one year. 
#salestsDec <- ts(salests[,2], frequency=52, start=c(2010, 1), end=c(2014,12))
#salestsDec <- decompose(salestsDec, type=c("additive"))
#plot(salestsDec)

```

Plot seasonally adjusted time series (Comments)

```{r}
# Check object type
class(clemCommentsDec1$seasonal)
class(clemCommentszDec)
# Both are 'ts' object, so we need to convert to numeric to be able to perform math operation of subtracting seasonal component from the original series. 


clemCommentszDec <- as.numeric(unlist(clemCommentszDec))
class(clemCommentszDec)
clemCommentsSeasonallyAdjusted <- clemCommentszDec - clemCommentsDec1$seasonal

#Plot seasonally adjusted time series
plot(clemCommentsSeasonallyAdjusted, xlab="Time", ylab="Comments mentions", main="Seasonally adjusted clementine Comment mentions")

#Combine data
clemCommentsFinal <- cbind(clemCommentsDec1$x,clemCommentsDec1$seasonal,clemCommentsSeasonallyAdjusted)

require(xlsx)
# Don't re-run, the file will append the current combined file. Write original price, seasonal component and seasonally adjusted values to en excel file

write.xlsx(clemCommentsFinal, file="/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clem_social_seasonally_adjusted_comments.xlsx", sheetName="sheet1", col.names=TRUE, row.names=FALSE, append=FALSE, showNA=TRUE)

```

Indexing can be an effective means of normalizing data to a common starting point and observing how variables change over time relative to each other.

Indexing: Modifying two or more numeric data series so that the resulting series start at the same value and change at the same rate as the unmodified series. 

Normalizing data for Platform Growth - The total volume of tweets / FB posts and so on for each of the social media platforms. In the case of the twitter data here, the search term was actually the total # of times the word "the" was tweeted (so as to provide a search term), but can be considered a proxy for total activity.

Read platform growth data 

```{r}
platform <- read.csv("/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/Platform_growth.csv")
```

Indexing platform growth data. There seems to be some noise in the beginning few weeks in data. Therefore, we index the data to 12/11/11 as the observations for each of the social media for that week appears to be fine. 

Excel: Indexed social media and platform growth data. Calculated the ratio of indexed data sources to obtain a pure number that can be used to compare with sales data. 

Next steps: 

- Analyze sales and social media data by plotting the trendlines for both of the time series. 




