---
title: "Predict food trend using social media data"
output: html_document
---

Objective is to predict total sales volume using social media data. 


```{r}
require(scales)
```


Read datasets for clementine sales and social media.  

```{r}
social <- read.csv("/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clementine_social.csv")

# Check summary statistics
#summary(social)
head(social[1:9])

sales <- read.csv("/Users/kevalshah/Keval_Backup/University/UChicago/Capstone/Data/clementine_sales.csv")

# Check summary statistics
#summary(sales)
head(sales)
```

Let's make some time series plots for social media mentions and sales data for Clementines. 

```{r}
# Social media mentions plot

# Date formatting 
analysis_date = as.Date(social[,1], "%m/%d/%y")
str(analysis_date)

# Create a variable for month to aggregate / sum all observations for a particular month 
social$month = as.Date(cut(analysis_date, breaks = "month"))
#social

total_mentions = social[,9]

# Create a dataset containing analysis date, month of analysis date and total social media mentions for plotting a observations aggregated at a monthly level. 
social1 <- data.frame(analysis_date, social$month, total_mentions)
#social1

require(ggplot2)
# Time series plot by monthly total observations
p <- ggplot(social1, aes(x=social.month, y=total_mentions)) 
p + geom_line() 
# add up all observations(total_mentions) for the month
p = p + stat_summary(fun.y = sum, geom = "line")
p = p + ggtitle ("Total clementine mentions across different social media outlets") 
p + xlab("Date") +  ylab ("Total mentions for Clementine")
#p + scale_x_date(labels = date_format("%y/%m"), breaks = "1 month")

# Plotting time series with curve / trendline 
#ggplot(socialts,aes(x=social.month, y=total_mentions)) + 
#  geom_point(aes(colour = total_mentions)) +
#  scale_colour_gradient2(low = "blue", mid = "green" , high = "red", midpoint = 4738 ) + 
#  geom_smooth(color = "red",size = 1) +
#  scale_y_continuous(limits = c(899,23695)) +
#  ggtitle ("Total clementine mentions across different social media outlets") +
#  xlab("Date") +  ylab ("Total mentions for Clementine")
   

```

```{r}
# Sales data plot 

# Date formatting 
period = as.Date(sales[,1], "%m/%d/%y")
str(period)

Volume = sales[,6]

# Create a variable for month to aggregate / sum all observations for a particular month 
sales$month = as.Date(cut(period, breaks = "month"))
#sales

# Bind period, sales month and volume into a data frame 
sales1 <- data.frame(period, sales$month, Volume)

# Now, let's plot time series of montly aggregated total Clementine sales volume
p <- ggplot(sales1, aes(x=sales.month, y=Volume))
p + geom_line() 
# add up all observations(total_mentions) for the month
p = p + stat_summary(fun.y = sum, geom = "line")
#p = p + scale_y_continuous(limits = c(211151,10017414))
p = p + ggtitle ("Total volume of clementine sales (in U.S)") 
p + xlab("Date") +  ylab ("Total Volume of Clementine sales")
```

Next steps: 

- Review time series textbook / literature for the right technique to deal with a seasonality:
detrending, differencing (average out)
- Understanding time series seasonal fluctuations or random fluctuations. 
- Decomposing the time series. There appears to be some seasonality in sales data. Look into smoothing, differencing and other time series techniques to control for seasonality. 
- Stationarity check. 

http://a-little-book-of-r-for-time-series.readthedocs.org/en/latest/src/timeseries.html 

Based on text readings: 

1. Decompose the time series and obtain trend, seasonal and random noise component from series. 
2. Remove seasonal component. The remaining time series should be trend and random noise. 
3. Regression modeling: Use dummy variables for each of months, maybe season as predictors and run the model. Note R-Squared. 



Decomposing the time series into trend, seasonal and irregular components. 

```{r}
# Columns required time and volume

#Using zoo object
require(zoo)
salests = sales[, c(1,6)]
salestsz = read.zoo(salests, format = "%m/%d/%y", index.column = 1)
View(salestsz)

#Decompose zoo object
salestszDec <- ts(salestsz, frequency=52, start=c(2010, 1), end=c(2014,12))
salestsDec1 <- decompose(salestszDec, type=c("additive"))
plot(salestsDec1)


# Frequency = 52 weeks in year. i.e 52 # of observations in one year. 
#salestsDec <- ts(salests[,2], frequency=52, start=c(2010, 1), end=c(2014,12))
#salestsDec <- decompose(salestsDec, type=c("additive"))
#plot(salestsDec)

```

Plot seasonally adjusted time series 

```{r}
# Check object type
class(salestsDec1$seasonal)
class(salestszDec)
# Both are 'ts' object, so we need to convert to numeric to be able to perform math operation of subtracting seasonal component from the original series. 


salestszDec <- as.numeric(unlist(salestszDec))
class(salestszDec)
salestszSeasonallyAdjusted <- salestszDec - salestsDec1$seasonal

#Plot seasonally adjusted time series
plot(salestszSeasonallyAdjusted, xlab="Time", ylab="Volume of sales", main="Seasonally adjusted clementine sales volume")
```




